Sequential and parallel algorithms for multidimensional multiextremal optimization based on the nested scheme of dimensionality reduction

### 2017-10-06 2017/ParOptBook/Chapter5.tex
Chapters 3 and 5 have been included. There is a question about cross-references between chapters.

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                             % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Parallel computations for the multidimensional multiextremal optimization problems with the dimensionality reduction using Peano curves}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Name of First Author and Name of Second Author}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\section{Section Heading}
Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.

>>>>

%%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\documentclass[graybox]{svmult}
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                            %% not available on your system
%%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
                             %% when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}
%\usepackage{amsmath}
%\usepackage{multirow}
%\newcommand{\extr}{extr}
%
%\makeindex             % used for the subject index
                       %% please use the style svind.ist with
                       %% your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Sequential and parallel algorithms for multidimensional multiextremal optimization based on the nested  scheme of dimensionality reduction}
\titlerunning{Nested  scheme of dimensionality reduction}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Vladimir A.Grishagin and Ruslan A.Israfilov}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Vladimir A.Grishagin \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia   \email{vagris@unn.ru}
\and Ruslan A.Israfilov \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia 
\email{ruslan@israfilov.com}}
%
\maketitle

\abstract{The dimensionality reduction schemes occupy an important place among the complexity reduction approaches in the multiextremal optimization problems. In the present chapter, one of such schemes, namely, so called nested optimization scheme is considered. This scheme allows one to replace solving a multidimensional problem by solving a family of one-dimensional subproblems nested recursively. Along with a general reduction scheme, its implementation in combination with various global search algorithms is considered. The nested scheme has a considerable potential of parallelization since various one-dimensional parallel optimization algorithms, in particular, the characteristical ones can be applied for solving the internal one-dimensional subproblems. The asynchronism of the parallel computations originating from the recursive nesting of the subproblems is an integral part of the scheme to be considered.}

\section{General nested optimization scheme}
\label{sec:5_1}
Let us consider a nonlinear programming problem in the form 
\begin{equation}
\label{eq:5_1}
\varphi(y)\rightarrow min,y\in Q\subseteq R^N,
\end{equation}
\begin{equation}
\label{eq:5_2}
Q=\{y\in D:g_j(y)\leq 0,\:1\leq j\leq m\},
\end{equation}
\begin{equation}
\label{eq:5_3}
D=\{y\in R^N:a_i\leq y\leq b_i,\:1\leq i\leq N\}.
\end{equation}

Assume that all constraint functions $g_j(y),\:1\leq j\leq m$,  are continuous and the domain $D$  is  bounded that provides the compactness of the feasible domain $Q$.

Let us introduce a continuous function $G(y)$  defined in the domain $D$  such that
\begin{equation}
\label{eq:5_4}
  \begin{cases}
    G(y)\leq 0, & y\in Q, \\
    G(y)>0, & y\notin Q.
  \end{cases}
\end{equation}

For example, the following functions can be taken as $G(y)$ :
\begin{equation}
\label{eq:5_5}
G(y)=\max\{g_1{y},\;g_2(y),\:\ldots\:,g_m(y)\}
\end{equation}
or
\begin{equation}
\label{eq:5_6}
G(y)=\max\{0;\:g_1{y},\;g_2(y),\:\ldots\:,g_m(y)\}.
\end{equation}
The latter function equals to zero at all points of the domain $Q$.

The following notations
\begin{equation}
\label{eq:5_7}
u_i=(y_1,\ldots,\:y_i),\;v_i=(y_{i+1},\ldots,y_N),
\end{equation}
allow one to write down the vector $y$ as a pair $y=(u_i,\;v_i)$  for $1\leq i\leq N$. Assume that $y=v_0$  if $i=0$  and $y=u_n$  for $i=N$.

Let us introduce  \textit{sections} of the domain $Q$ as
\begin{equation}
\label{eq:5_8}
S_1=Q,\ S_{i+1}(u_i)=\{v_i\in R^{N-i}:(u_i,v_i)\in Q\},\:1\leq i\leq N-1,
\end{equation}
and \textit{projections}
\begin{equation}
\label{eq:5_9}
\Pi_{i+1}(u_i)=\{y_{i+1}\in R^1:\exists (u_{i+1},v_{i+1})\in S_{i+1}(u_i)\},\:1\leq i\leq N-1,
\end{equation}
of the sections $S_{i+1}(u_i)$ onto the axis $y_{i+1}$.

Let us define recursively a family of functions
\begin{equation}
\label{eq:5_10}
  \begin{cases}
    G^N(y)\equiv G(y), \\
    G^i(u_i)=\min\{G^{i+1}(u_i,y_{i+1}:y_{i+1}\in [a_{i+1},b_{i+1}]\},\;1\leq i\leq N-1,
  \end{cases}
\end{equation}
over corresponding projections 
\begin{equation}
\label{eq:5_11}
D_i=\{u_i\in R^i:y_j\in[a_j,b_j],\:1\leq j\leq i\}
\end{equation}
of the domain $D$  from (\ref{eq:5_3}) onto the coordinate axes $y_1,\ldots,y_i$. Moreover, by definition, $D_N=D$.

Because of the continuity of the function $G^N(y)\equiv G(y)$ and of the compactness of the domain $D$ the function  $G^{N-1}(u_{N-1})$ exists and is continuous in  $D_{N-1}$. This fact causes the existence and the continuity of the function  $G^{N-2}(u_{N-2})$ in  $D_{N-2}$ and further determines the existence and the continuity of all functions from family (\ref{eq:5_10}) in corresponding domains (\ref{eq:5_11}).

The following lemma takes place.
\begin{lemma} 
\label{lem:5_1}
\begin{equation}
\label{eq:5_12}
G^i(u_i)=\min\{G(u_i,v_i):y_j\in [a_j,b_j],\:i+1\leq j\leq N\}.
\end{equation}
\end{lemma}
\begin{proof}
As a consequence of (\ref{eq:5_10}) the assertion (\ref{eq:5_12}) is equivalent to
\begin{equation}
\label{eq:5_13}
\min_{y_{i+1}\in [a_{i+1},b_{i+1}]}\ldots \min_{y_N\in [a_N,b_N]}G(u_i,v_i)=\min\{G(u_i,v_i):y_j\in [a_j,b_j],i+1\leq j\leq N\}.
\end{equation}

Because of the continuity of $G(y)$, for any $u_i\in D_i$ there exists $\bar{v}_i=(\bar{y}_{i+1},\ldots,\bar{y}_N)$ such that $\bar{y}_j\in [a_j,b_j],\:i+1\leq j\leq N,$ and
\begin{displaymath}
\min_{y_{i+1}\in [a_{i+1},b_{i+1}]}\ldots \min_{y_N\in [a_N,b_N]}G(u_i,v_i)=G(u_i,\bar{v}_i)
\end{displaymath}
from where it follows that the left-hand side of the equality  (\ref{eq:5_13}) is greater or equal to the right-hand side.

Now the subject of our interest is the reverse inequality. Because of the continuity of $G(y)$ and of the compactness of the domain $D$  there exists a vector $v_i^*=(y_{i+1}^*,\ldots ,y_N^*)$  such that 
\begin{displaymath}
G(u_i,v_i^*)=\min\{G(u_i,v_i):y_j\in [a_j,b_j],\:i+1\leq j\leq N\}.
\end{displaymath}

According to (\ref{eq:5_10}), we have
\begin{displaymath}
G^{N-1}(u_i,y_{i+1}^*,\ldots ,y_{N-1}^*)=\min\{G(u_i,y_{i+1}^*,\ldots,y_{N-1}^*,y_N):y_N\in [a_N,b_N]\}\leq G(u_i,v_i^*),
\end{displaymath}
\begin{displaymath}
G^{N-2}(u_i,y_{i+1}^*,\ldots ,y_{N-2}^*)=\min\{G^{N-1}(u_i,y_{i+1}^*,\ldots,y_{N-2}^*,y_{N-1}):y_{N-1}\in [a_{N-1},b_{N-1}]\}\leq 
\end{displaymath}
\begin{displaymath}
\leq G^{N-1}(u_i,y_{i+1}^*,\ldots,y_{N-2}^*,y_{N-1}^*)\leq G(u_i,v_i^*),
\end{displaymath}
$\ldots$
\begin{displaymath}
G^i(u_i)=\min\{G^{i+1}(u_i,y_{i+1}):y_{i+1}\in [a_{i+1},b_{i+1}]\}\leq G^{i+1}(u_i,y_{i+1}^*)\leq G(u_i,v_i^*).
\end{displaymath}
This chain of inequalities finalizes the proof.
\end{proof}

Let us introduce the projections 
\begin{equation}
\label{eq:5_14}
Q_i=\{u_i\in R^i:\exists (u_i,v_i)\in Q\},\:1\leq i\leq N,
\end{equation}
of the domain $Q$ onto the coordinate axes $y_1,\ldots,y_i$.
\begin{lemma} 
\label{lem:5_2}
Projection (\ref{eq:5_14}) can be presented in the equivalent form 
\begin{equation}
\label{eq:5_15}
Q_i=\{u_i\in R^i:G^i(u_i)\leq 0\}.
\end{equation}
\end{lemma}
\begin{proof}
Let  (\ref{eq:5_14}) be fulfilled, i.e., for some $u_i$  there exists $v_i^*$  such that $(u_i,v_i^*)\in Q$. Then, $G(u_i,v_i^*)\leq 0$ , i.e., as a consequence of Lemma \ref{lem:5_1},
\begin{displaymath}
G^i(u_i)=\min\{G(u_i,v_i):y_j\in[a_j,b_j],\:i+1\leq j\leq N\}\leq G(u_i,v_i^*)\leq 0,
\end{displaymath}
and (\ref{eq:5_15}) is true.

Now assume that $G^i(u_i)\leq 0$  for some $u_i\in Q_i$. However, according to Lemma \ref{lem:5_1}, a vector $v_i^*$  exists such that $G^i(u_i)=G(u_i,v_i^*)\leq 0$ , i.e., $(u_i,v_i^*)\in Q$.

Lemma has been proved.
\end{proof}
\begin{lemma} 
\label{lem:5_3}
The definition of the projection (\ref{eq:5_9}) is equivalent to the form 
\begin{equation}
\label{eq:5_16}
\Pi_{i+1}(u_i)=\{y_{i+1}\in [a_{i+1},b_{i+1}]:G^{i+1}(u_i,y_{i+1})\leq 0\}.
\end{equation}
\end{lemma}
\begin{proof}
First of all, note that because of $Q\subseteq D$, for all $y\in Q$ the double inequalities $a_j\leq y_j\leq b_j$,  , hold necessarily. Let $y_{i+1}$  be such that $G^{i+1}(u_i,y_{i+1})\leq 0$ . Then there exists $v_{i+1}^*$  such that $G(u_i,y_{i+1},v_{i+1}^*)=G^{i+1}(u_i,y_{i+1})\leq 0$, i.e., $(y_{i+1}, v_{i+1}^*)\in S_{i+1}(u_i)$. Consequently, $y_{i+1}$  belongs to the projection $\Pi_{i+1}(u_i)$  in the sense of the definition (\ref{eq:5_9}).

Next, for any $y_{i+1}$  satisfying (\ref{eq:5_9}) there exists $v_{i+1}^*$  such that $(y_{i+1},v_{i+1}^*)\in S_{i+1}(u_i)$. Then, the vector $(u_i,y_{i+1},v_{i+1}^*)\in Q$, i.e., $G^{i+1}(u_i,y_{i+1})\leq G(u_i,y_{i+1},v_{i+1}^*)\leq 0$.

The equivalence of (\ref{eq:5_9}) and (\ref{eq:5_16})  has been proved.
\end{proof}
Now consider  a continuous function $\varphi(y)$. Setting $\varphi^N(y)\equiv \varphi(y)$ by definition, let us construct a family of functions 
\begin{equation}
\label{eq:5_17}
\varphi^i(u_i)=\min\{\varphi^{i+1}(u_i,y_{i+1}):y_{i+1}\in \Pi_{i+1}(u_i)\},\:1\leq i\leq N-1,
\end{equation}
defined over the corresponding projections $Q_i$  from (\ref{eq:5_14}). Then, according to \cite{5_CarrHowe} and \cite{5_StrSergMon2000} the basic relation of the nested optimization scheme 
\begin{equation}
\label{eq:5_18}
\min_{y\in Q}\varphi(y)=\min_{y_1\in \Pi_1}\min_{y_2\in \Pi_2(u_1)}\ldots\min_{y_N\in \Pi_N(u_{N-1})}\varphi(y).
\end{equation}
takes place.

As it follows from (\ref{eq:5_18}), in order to solve the problem (\ref{eq:5_1})--(\ref{eq:5_3}) it is sufficient to solve the one-dimensional problem 
\begin{equation}
\label{eq:5_19}
\varphi^1(y_1)\rightarrow \min,\:y_1\in \Pi_1\subseteq R^1,
\end{equation}
\begin{displaymath}
\Pi_1=\{y_1\in [a_1,b_1]:G^1(y_1)\leq 0\}.
\end{displaymath}

According to (\ref{eq:5_17}), each calculation of the function $\varphi^1(y_1)$  at some fixed point $y_1\in\Pi_1$   requires solving  the one-dimensional problem 
\begin{displaymath}
\varphi^2(y_1,y_2)\rightarrow \min,\:y_2\in \Pi_2(y_1)\subseteq R^1,
\end{displaymath}
\begin{displaymath}
\Pi_2(y_1)=\{y_2\in [a_2,b_2]:G^2(y_1,y_2)\leq 0\}.
\end{displaymath}
This problem is a one-dimensional minimization problem with respect to $y_2$  since $y_1$  is fixed.

In turn, each calculation of the function $\varphi^2(y_1,y_2)$  with fixed $y_1,y_2$ generates solving the one-dimensional problem
\begin{displaymath}
\varphi^3(u_2,y_3)\rightarrow \min,\:y_3\in \Pi_3(u_2)\subseteq R^1,
\end{displaymath}
etc., up to solving the univariate problem

\begin{equation}
\label{eq:5_20}
\varphi^N(u_{N-1},y_N)\rightarrow \min,\:y_N\in \Pi_N(u_{N-1})\subseteq R^1,
\end{equation}
where $u_{N-1}$  is fixed.

Thus, solving the problem (\ref{eq:5_1})--(\ref{eq:5_3}) can be reduced to solving a family of «nested» one-dimensional subproblems 
\begin{equation}
\label{eq:5_21}
\varphi^i(u_{i-1},y_i)\rightarrow \min,\:y_i\in \Pi_i(u_{i-1}),
\end{equation}
where the fixed vector $u_{i-1}\in Q_{i-1}$.

Solving initial multidimensional problem (\ref{eq:5_1})--(\ref{eq:5_3})  through solving the system of the one-dimensional subproblems (\ref{eq:5_21}) is called \textit{the nested optimization scheme}.

On the base of the nested scheme (\ref{eq:5_18}) many multidimensional optimization algorithms have been designed (see, for example, \cite{5_CarrHowe, 5_Evtushenko, 5_GerGriIsr, 5_GriIsrAIP, 5_GriIsrCEUR, 5_GrishaginStrongin_EnginCybernetics, 5_Piyavskij, 5_SergGriJCAA, 5_ShiOlaf, 5_StrSergMon2000, 5_vanDam}). This scheme has essential potential for constructing new optimization techniques by means of combining with different one-dimensional algorithms for solving the subproblems (\ref{eq:5_21}) (for example, characteristical methods \cite{5_GrishaginSergeyevStrongin} including information    algorithms \cite{5_SergeyevLocTun, 5_Sergeyev1998, 5_SergeyevGrishaginOMS, 5_SergMukhKvasLera, 5_StrMonRus, 5_Str1973, 5_StrMarkin, 5_StrSergMon2000}, methods by Piyavskij and Shubert \cite{5_Piyavskij, 5_Shubert} and Kushner \cite{5_Kushner}, Bayesian algorithms \cite{5_Locatelli, 5_Mockus1980, 5_Mockus1988, 5_MockusEddy_et_al, 5_Zilinskas1975, 5_Zilinskas1981, 5_Zilinskas1985}, etc.)

As an illustration, let us consider an example of the maximization of a two-dimensional multiextremal function from the class (\ref{eq:3_518}) using the nested optimization scheme. The surface and level curves of the objective function are shown in Fig. 5.1 and in Fig. \ref{fig:5_2}, respectively. The coordinates of the search trials of the nested scheme are shown in Fig. 5.2 also. For this function the global maximum value 13.51 is located at the point (0.603, 0.408). The method AGS (\ref{eq:2_38})--(\ref{eq:2_40}) was employed for solving the one-dimension subproblems (\ref{eq:5_21}).
\begin{figure}[t]
\centering
\includegraphics[width=1.1\linewidth]{figures/figure_5_2.pdf}
\caption{Level curves and trial points}
\label{fig:5_2}    
\end{figure}

The nested scheme (\ref{eq:5_18})--(\ref{eq:5_21}) may be generalized in the following way. Let us consider for simplicity a problem (\ref{eq:5_1}) without functional constraints, i.e., $Q=D$, and present the feasible domain  as the direct product of subdomains with smaller dimensions
\begin{displaymath}
D=D_1\times D_2\times\:\ldots\: D_s,\:1\leq s<N,
\end{displaymath}
where the subdomains $D_i,\:1\leq i\leq s,$ are hyperparallelepipeds (\ref{eq:5_3}) of corresponding dimensions $d_i$ such that $1\leq d_i<N$ and $\sum_{i=1}^s{d_i}=N$.

In this case (\ref{eq:5_18}) can be transformed to the form
\begin{displaymath}
\min_{y\in Q}\varphi(y)=\min_{z_1\in D_1}\min_{z_2\in D_2}\ldots\min_{z_s\in D_s}\varphi(y),
\end{displaymath}
where $z_1=u_{\sigma(1)},\:z_i=(y_{\sigma{(i-1)}}+1,\ldots, y_{\sigma(i)}),\:2\leq i\leq s,$ and $\sigma(j)=\sum_{i=1}^j{d_i},\:1\leq j\leq s$.

Then, instead of the family (\ref{eq:5_21}) we can solve the set of nested subproblems 
\begin{displaymath}
\varphi(u_{\sigma(i-1)},z_i)\rightarrow \min,\;z_i\in D_i,
\end{displaymath}
which can be multidimensional with lesser dimensions. Such approach was developed (see \cite{5_BarkGer2014, 5_BarkGerLeb, 5_BarkLeb, 5_SysoyevBarkGerLeb}) in combination of the optimization algorithms based on the Peano space-filling curves for solving the nested multidimensional subproblems.

Moreover, in the family (\ref{eq:5_21}) for several functions $\varphi^i(u_i)$ one can use the operation of taking maximum instead of minimum. In this case it is possible to calculate complex minmax (or maxmin) expressions.

\section{Properties of one-dimensional subproblems in the nested optimization scheme}
\label{sec:5_2}
While analyzing univariate subproblems (\ref{eq:5_21}) of the nested scheme, two main problems arise:
\begin{description} [a)]
\item [a)] {the necessity to construct the feasible domains $\Pi_i(u_{i-1})$  for the one-dimensional search;}
\item [b)] {it is required to provide the minimization of the one-dimensional functions $\varphi^i(u_{i-1},y_i)$  in the domains  $\Pi_i(u_{i-1})$.}
\end{description}

The structure and the complexity of the projections $\Pi_i(u_{i-1})$  are completely determined by the complexity of the multidimensional feasible domain $Q$. The complexity of the second problem depends on the characteristics of the functions $\varphi^i(u_i)$ . These characteristics are influenced by the properties of the objective function $\varphi(y)$ as well as by the features of the search domain $Q$ defined, in turn, by constraints (\ref{eq:5_2}) and (\ref{eq:5_3}).

\subsection {Structure of the feasible domains of the one-dimensional search}
\label{subsec:5_2_1}
The results of Lemma \ref{lem:5_3}, which has established the equivalence of Definition (\ref{eq:5_9}) and representation (\ref{eq:5_16}) enable to analyze the structure of the domains $\Pi_i(u_{i-1})$. In fact, (\ref{eq:5_16}) provides a constructive apparatus for building the domains $\Pi_i(u_{i-1})$  by means of finding the domains of the non-positivity for the functions $G^i(u_{i-1},y_i)$.

Since the function $G(y)$ is assumed to be continuous in the domain $D$, the functions $G^i(u_i)$ are continuous with respect to $u_i\in D_i$  from (\ref{eq:5_11}) and, therefore, with regard to $y_i\in [a_i,b_i]$  as well. Then, at fixed $u_{i-1}\in D_{i-1}$  each one-dimensional problem (\ref{eq:5_21}) can be rewritten in a unified form
\begin{equation}
\label{eq:5_22}
  \begin{cases}
    \bar{\varphi}(x)\rightarrow\min,\;x\in\bar{Q}\subseteq R^1, \\
    \bar{Q}=\{x\in[a,b]:g(x)\leq 0\},
  \end{cases}
\end{equation}
where the function $g(x)$ is continuous.

The continuity of the constraint $g(x)$ allows one to state that the feasible domain $\bar{Q}$  can be presented as a system of intervals
\begin{equation}
\label{eq:5_23}
\bar{Q}=\bigcup_{j=1}^q{[a^j,b^j]},
\end{equation}
in each of which the function $g(x)$ is non-positive.

As an example, let us consider the possible cases of the behavior of a continuous function generating the domains of the non-positivity presented in Fig. \ref{fig:5_3}. The corresponding intervals  are marked with  the signs $\times$  on the abscissa axis including the points of tangency. The most right-side interval marked with $\times$ corresponds to the case when the function equals to zero over the interval of non-zero length. 
\begin{figure}[t]
\centering
\includegraphics[width=1.1\linewidth]{figures/figure_5_3.pdf}
\caption{Domains of the non-positivity of a continuous function}
\label{fig:5_3}    
\end{figure}
In the system (\ref{eq:5_23}) the number of intervals $q$ can be infinite. As an example of such situation one can consider within the closed interval $[0,1]$ the function
\begin{displaymath}
g(x)=
  \begin{cases}
    x\sin(\frac{1}{x}), & x>0,\\
    0, & x=0.
  \end{cases}
\end{displaymath}

Thus, in the case of a continuous function $G(y)$, the projection $\Pi_i(u_{i-1}$ determined as  (\ref{eq:5_16}) is a set of the kind (\ref{eq:5_23}), i.e.,
\begin{equation}
\label{eq:5_24}
\Pi_i(u_{i-1})=\bigcup_{j=1}^{q_i}{[a_i^j,b_i^j]},
\end{equation}
where, in general, the number of the intervals $q_i$  and their bounds $a_i^j$, $b_i^j$, $1\leq j\leq q_i,$ depend on the vector $u_{i-1}$ , i.e.,
\begin{equation}
\label{eq:5_25}
q_i=q_i(u_{i-1}),\:a_i^j=a_i^j(u_{i-1}),\:b_i^j=b_i^j(u_{i-1}).
\end{equation}

If the domain $Q$  is such that it is possible to obtain the explicit (analytical) expressions for the values $q_i,a_i^j,b_i^j$   as  functions of $u_{i-1})\in Q_{i-1}$  for all $1\leq i\leq N$, then the domain $Q$ is called \textit{the domain with the computable bounds}. In particular, it is possible if all the roots of the functions $G^i(u_{i-1},y_i)$  with regard to the variables $y_i$  can be analytically found.
\begin{example}
\label {exam:5_1}
As the problem to be investigated  the problem (\ref{eq:5_1})--(\ref{eq:5_3}) with 
\begin{equation}
\label{eq:5_26}
  \begin{cases}
    \varphi(y)=y_1^2+y_2^2,\\
   Q=\{y\in D:(y_1-1)^2+(y_2-1)^2-1\leq 0\}, \\
	D=\{y\in R^2:0\leq y_1,y_2\leq 2\}
  \end{cases}
\end{equation}
is considered. The feasible domain is marked with gray in Fig. \ref{fig:5_4}. Also, the level curves of the objective function are shown there. In this problem, the function $G^2(y)=(y_1-1)^2+(y_2-1)^2-1$. It is evident that the function  $G^1(y_1)=\min\{G^2(y_1,y_2):y_2\in [0,\;2]\}=(y_1-1)^2-1$. 
\begin{figure}[t]
\centering
\includegraphics[width=1.1\linewidth]{figures/figure_5_4.pdf}
\caption{Feasible domain and level curves of the objective function}
\label{fig:5_4}    
\end{figure}

According to (\ref{eq:5_16}), the domains where the functions $G^1(y_1)$  for $y_1$  and $G^2(y_1,y_2)$  for $y_2$  are not positive determine the projections $\Pi_1$  and $\Pi_2(y_1)$ , respectively. The domain bounds are the roots of these functions in the interval $[0,2]$. For the function $G^1(y_1)$  these roots are $0$ and $2$. Therefore, $\Pi_1=[0,\:2]$.

The function $G^2(y)=(y_2-1)^2-\alpha^2$, where $\alpha=\sqrt{1-(y_1-1)^2}\leq 1$,  has the roots $1\pm \alpha$ within the interval $[0,2]$ and  it is not positive between these roots. As a result, 
\begin{equation}
\label{eq:5_27}
\Pi_2(y_1)=[1-\sqrt{1-(y_1-1)^2},1+\sqrt{1-(y_1-1)^2}].
\end{equation}

Thus, we can write down the bounds of projections (\ref{eq:5_24}) in an explicit form. Thereby, the domain $Q$ from (\ref{eq:5_26}) is a domain with computable bounds (\ref{eq:5_25}).

The function $\varphi(y)=y_1^2+y_2^2$  is an increasing one for $y_2$  in the interval $[0,2]$. Consequently, it reaches its maximum at the point $1-\alpha$  in the domain (\ref{eq:5_27}), therefore,
\begin{displaymath}
\varphi^1(y_1)=y_1^2+(1-\sqrt{1-(y_1-1)^2})^2=1+2y_1-2\sqrt{1-(y_1-1)^2}.
\end{displaymath}

The first derivative $(\varphi^1(y_1))'=2+\frac{2(y_1-1)}{\sqrt{1-(y_1-1)^2}}$ has a unique root  $y_1^*=1-\frac{1}{\sqrt{2}}$ in the interval  $[0,2]$. 
In addition, the second derivative $(\varphi^1(y_1))''=\frac{2}{(1-(y_1-1)^2)^{3/2}}>0$, therefore, the point $y_1^*=1-\frac{1}{\sqrt{2}}$  gives the minimum value of the function $\varphi^1(y_1)$  equal to $3-2\sqrt{2}$. This value is the sought minimum of the function $\varphi(y)$ in the domain $Q$.

In order to obtain the coordinate $y_2^*$, which determines the minimum point along with $y_1^*$, let us consider the function $\varphi^2(y_1^*,y_2)$ and find its minimum in the domain $\Pi_2(y_1^*)=[1-\frac{1}{\sqrt{2}},1+\frac{1}{\sqrt{2}}]$, which is obviously achieved at the point $y_2^*=1-\frac{1}{\sqrt{2}}$.

So, the vector $y^*=(1-\frac{1}{\sqrt{2}},1-\frac{1}{\sqrt{2}})$  providing the minimum value $\varphi^*=3-2\sqrt{2}$ of the objective function is the solution of the multidimensional problem (\ref{eq:5_26}). The optimum point is marked with a dark cross in Fig. \ref{fig:5_4}. 
\end{example}

In the example considered above, we have constructed the one-dimensional search domain bounds  analytically by finding the domains of the non-positivity of the corresponding functions $G^i(u_i)$ . At the same time, one can specify a more visual «geometrical» method for the construction of the projections $\Pi_{i+1}(u_i)$. This method follows from definitions (\ref{eq:5_8}) and (\ref{eq:5_9}) and consists in the construction of the sections of the domain $Q$  by planes $u_i=const$. After that, it is necessary to find the bounds of these sections for the coordinate $y_{i+1}$.

In this regard, the following is noteworthy. The computation of the global minimum according to (\ref{eq:5_18}) is analogous to the procedure of computation of a multidimensional integral of the function $\varphi(y)$ over the domain $Q$ by means of the reduction to computing  the recurring (nested) one-dimensional integrals. Here, the domains of the one-dimensional integration are just the corresponding projections $\Pi_i(u_{i-1})$.

As an illustration the example \ref{exam:5_2} can be taken. 
\begin{example}
\label{exam:5_2}
Assume that the domain of the optimization (or of the integration) is defined as
\begin{displaymath}
Q=\{y\in R^2:-4\leq y_1,y_2\leq 4, y_1^2+y_2^2\leq 4,y_1^2+y_2^2\geq 1\}.
\end{displaymath}
\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{figures/figure_5_5.pdf}
\caption{Sections and projections}
\label{fig:5_5}    
\end{figure}
In Fig. \ref{fig:5_5} this domain is marked with gray. The left-side line $y_1=const$  generates the section $S_1(y_1)$  as the intersection of the line with the domain $Q$, and the projection  of $S_1(y_1)$  onto the axis $y_2$  determines the projection $\Pi_2(y_1)$ . This projection and the projection $\Pi_1$  are marked with the bold lines in Fig. \ref{fig:5_5}.

These geometrical considerations allow constructing the necessary projections of the problem as 
\begin{displaymath}
\Pi_1=[-2,\:2],
\end{displaymath}
\begin{displaymath}
\Pi_2(y_1)=
  \begin{cases}
    [-\sqrt{4-y_1^2}, \sqrt{4-y_1^2}],\ y_1\in [-2,\:-1]\bigcup [1,\:2],  \\
    [-\sqrt{4-y_1^2}, -\sqrt{1-y_1^2}]\bigcup [\sqrt{1-y_1^2}, \sqrt{4-y_1^2}],y_1\in [-1,\:1].                      	
  \end{cases}
\end{displaymath}
\end{example}

However, finding all roots of a continuous function is a difficult problem, unsolvable analytically, as a rule. In this case, one can try to find these roots numerically. Let us consider, for example, typical problem (\ref{eq:5_22}). In order to find all roots of a continuous function $g(x)$ one can propose  numerical solving  the equivalent optimization problem 
\begin{displaymath}
\left|g(x)\right|\rightarrow \min,x\in [a,b],
\end{displaymath}
which roots of $g(x)$ are the global minima points in. To solve this problem, the characteristical global search algorithms providing the convergence to all global minimum points might be applied.

Another approach to accounting the constraints in the optimization problems (the index method), which does not require to solve any auxiliary problems  is considered in Chapter 4.

The case $Q=D$  when the constraint functions (\ref{eq:5_2}) are absent is an important particular case of the problem (\ref{eq:5_1})--(\ref{eq:5_3}). In this case, $G(y)\equiv 0$ in the domain $D$, and it follows from (\ref{eq:5_12}) that $G^i(u_i)\equiv 0$,$1\leq i\leq N$. Then, according to (\ref{eq:5_16}), 
\begin{displaymath}
\Pi_i(u_{i-1})=[a_i,b_i],
\end{displaymath}
where $a_i,b_i$ are constants.

The optimization domains (\ref{eq:5_2}) can generate projections $\Pi_i(u_{i-1})$  as a single interval for some particular properties of the constraint functions as well. These properties are the convexity of the functions $g_j(y),\:1\leq j\leq m$, providing the convexity of the feasible domain $Q$ and the monotonous unimodality of the constraint functions \cite{5_GrishaginStrongin_EnginCybernetics}, which is more general than the concept of convexity.  In general case, monotonous unimodal constraints can generate non-convex domains $Q$.
